### 2024-03-11 paper/paper.qmd
Initial commit

---
title: "Operating Budget Program Summary By Expensiture Category"
subtitle: "Murrumbidgee Paper"
execute:
  echo: false
  message: false
  warning: false
author: 
  - Joe Zhou
thanks: "Code and data are available at: LINK."
date: today
date-format: long
mainfont: "Hiragino Sans GB"
format:
  pdf:
    documentclass: scrartcl
    geometry:
      - top=30mm
      - left=20mm
      - heightrounded
    fontfamily: libertinus
    colorlinks: true
    toc: true
classoption: abstract
abstract: | 
  text
include_in_header:
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
---

```{r}
library(dplyr)
library(modelsummary)
library(tidyverse)
library(ggplot2)
library(rstanarm)
library(here)
library(arrow)
library(tidybayes)
library(tidyverse)
library(marginaleffects)
library(readxl)
library(httr)
```

```{r}
#| output: false

```

# Introduction

# Data

# Model

# Results

# Discussion

\newpage

# References
@citeR